---
title: "Exploratory Data Analysis: Patients vs. Controls"
author: "Natalia Lara & Juan Martinez"
date: "3/16/2023"
output:
  html_document:
    toc: true
    number_sections: true
    toc_depth: 2
    toc_float:
      collapsed: true
      smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(tidyverse)
library(tidyr)
library(stringr)
library(readr)
library(kableExtra)
library(ggdendro)
library(plotly)
```

# Introduction

This document presents some preliminary analysis of the mobility dataset gathered in 2019 at the Military Hospital. 

# Data 

Random experiment. Two main groups: manual and optical system.

* Schobert: goihgosjg


* Flexion:


rotación

## Patients

```{r data1}
patients <- read_csv('Data/patients.csv')
patients$NUMERO <- as_factor(patients$NUMERO)
patients$ID <- as_factor(patients$ID)
patients$BIOLOGICO <- as_factor(patients$BIOLOGICO)
patients$TIPO <- as_factor(patients$TIPO)

sep_bp_patients <- str_split(patients$PRESION_ARTERIAL, '/')

SYSTOLIC = DIASTOLIC = 0

for(i in (1:nrow(patients))){
  SYSTOLIC[i] = as.numeric(sep_bp_patients[[i]][1])
  DIASTOLIC[i] = as.numeric(sep_bp_patients[[i]][2])
}

patients$SYSTOLIC <- SYSTOLIC
patients$DIASTOLIC <- DIASTOLIC

patients %>%
  kbl(booktabs = T) %>%
  kable_material(c('hover', 'striped'), full_width = F) %>%
  scroll_box(width = "100%", height = "250px")
```

## Controls

```{r data2}
controls <- read_csv('Data/controls.csv')

controls %>%
  kbl(booktabs = T) %>%
  kable_material(c('hover', 'striped'), full_width = F) %>%
  scroll_box(width = "100%", height = "250px")
```


# Differences

We compare the differences between the reported measurements from the manual and optical system. This comparison was made for each variable. For instance, we compared the results from the cervial rotation using the manual instrument and optical system.

## Rotacion Cervical (Grados)


```{r}
patients %>%
  ggplot(aes(x = ROTA_CERV_MANUAL_DERECHA)) +
  geom_density(aes(fill = 'Manual', colour = 'Manual'), alpha = 0.2) +
  geom_density(aes(x = ROTA_CERV_LAB_DERECHA, colour = 'Laboratory',
                   fill = 'Laboratory'), alpha = 0.2) +
  xlab('Rotacion Cervical Derecha (Grados)') +
  theme(legend.position = 'bottom')
```

```{r}
patients %>%
  ggplot(aes(x = ROTA_CERV_MANUAL_IZQUIERDA)) +
  geom_density(aes(fill = 'Manual', colour = 'Manual'), alpha = 0.2) +
  geom_density(aes(x = ROTA_CERV_LAB_IZQUIERDA, colour = 'Laboratory',
                   fill = 'Laboratory'), alpha = 0.2) +
  xlab('Rotacion Cervical Derecha (Grados)') +
  theme(legend.position = 'bottom')
```


## Flexion Frontal Cervical (Grados)


```{r}
fig3 <- patients %>%
  ggplot(aes(x = FLEX_FRONT_CERV_MANUAL_GRAD_FLEXION)) +
  geom_density(aes(fill = 'Manual', colour = 'Manual'), alpha = 0.2) +
  geom_density(aes(x = FLEX_FRONT_CERV_LAB_GRAD_FLEXION, colour = 'Laboratory',
                   fill = 'Laboratory'), alpha = 0.2) +
  xlab('Flexion Frontal Cervical (Grados)') +
  theme(legend.position = 'bottom')

ggplotly(fig3, tooltip = "all")
```


```{r}
fig4 <- patients %>%
  ggplot(aes(x = FLEX_FRONT_CERV_MANUAL_GRAD_EXTESION)) +
  geom_density(aes(fill = 'Manual', colour = 'Manual'), alpha = 0.2) +
  geom_density(aes(x = FLEX_FRONT_CERV_LAB_GRAD_EXTESION, colour = 'Laboratory',
                   fill = 'Laboratory'), alpha = 0.2) +
  xlab('Flexion Frontal Cervical (Grados)') +
  theme(legend.position = 'bottom')

ggplotly(fig4, tooltip = "all")
```


## Dissimilarity

```{r}
diss_manual <- dist(patients[,c(8,9,12,13,16,17,20, 22, 23,26)])
hc_patients_manual <- hclust(diss_manual)

plot(hc_patients_manual, xlab = 'Clusters - Manual')
```


```{r}
ggdendrogram(hc_patients_manual)
```

# Conclusions

The results suggests that there are significant differences  between the approaches (manual and optical system). Clearly, the optical system is more accurate than the manual approach and its results can be set as a *gold* standard. Also, these results suggest that even well trained personnel can report biased results.


# Appendix {-}

This section contains analysis or functions that can be used in the analysis. 

## BASMI scores

The BASMI score is a standard method to assess the mobility of patients with gait disorders. 

```{r}
basmi_score <- function(FLL, Schober, DI, RC){
  
  # Lateral lumbar Flexion score 
  if(FLL > 10){
    score_FLL=1
  }
  else if ((FLL >= 5) & (FLL<=10)){
    score_FLL=2
  }
  else{
    score_FLL=3
  }
  
  # Schober score
  
  if(Schober > 4){
    score_Schober=1
  }
  else if ((Schober >= 2) & (Schober <=4)){
    score_Schober=2
  }
  else{
    score_Schober=3
  }
  
  # Intermalleolar distance score
  
  if(DI > 100){
    score_DI=1
  }
  else if ((DI >= 70) & (DI <=100)){
    score_DI=2
  }
  else{
    score_DI=3
  }
  
  # Cervical rotation score
  
  if(RC > 70){
    score_RC=1
  }
  else if ((RC >= 20) & (RC <=70)){
    score_RC=2
  }
  else{
    score_RC=3
  }
  return(score_FLL+score_Schober+score_DI+score_RC)
}
basmi_score(10,19.5,28.3,50)
```
